---
title: "Machine Learning Course Project"
author: "Dan Killian"
date: "September 14, 2015"
output: html_document
---
```{r}
setwd("C:/Users/Dan/Dropbox/personal/Coursera/Data science certificate/Machine learning/course project/")

library(AppliedPredictiveModeling)
library(caret)
library(dplyr)
library(tidyr)
library(lubridate)
library(reshape2)
library(kernlab)
library(rattle)
library(ElemStatLearn)
library(ISLR)
library(stargazer)

test <- read.csv("pml-testing.csv")
train <- read.csv("pml-training.csv")
trainNames <- data.frame(names(train))
train[1:5,1:10]
str(train)
table(train$classe)

names(train)[1:10]

table(train$classe)
table(train$user_name)

train2 <- select(train, user_name, roll_belt:classe)
test2 <- select(test, user_name, roll_belt:problem_id)

tail(names(train))
tail(names(test))
```
Convert factor variables
```{r}
train <- apply(train[,12:17],2,as.numeric(as.character))

train$kurtosis_roll_belt <- as.numeric(as.character(train$kurtosis_roll_belt))

```

Let's try PCA. 

```{r}
procPCA <- preProcess(train[,8:159], method="pca", thresh=.8)
str(procPCA)

classe <- predict(procPCA, newdata=test[,8:159])
                  


procModFit <- train(classe~., method="rpart", data=)

```

Tree

```{r cache=TRUE}

modFit <- train(classe ~. , method="rpart", data=train2)
print(modFit$finalModel)
names(modFit)

plot(modFit$finalModel, uniform=T, main="Classification Tree")
text(modFit$finalModel, use.n=TRUE, all=TRUE, cex=.8)

fancyRpartPlot(modFit$finalModel)

confusionMatrix(t$classe, predict(modFit, newdata=test))
?confusionMatrix

res <- predict(modFit, newdata=test)
?predict

predictions <- predict(modelFit, newdata=testing)
predictions


```


```{r}

modFit2 <- train(classe ~. , method="rpart", data=train2)
print(modFit2$finalModel)
names(modFit2)

plot(modFit2$finalModel, uniform=T, main="Classification Tree")
text(modFit2$finalModel, use.n=TRUE, all=TRUE, cex=.8)

fancyRpartPlot(modFit2$finalModel)

confusionMatrix(t$classe, predict(modFit, newdata=test))
?confusionMatrix

res <- predict(modFit2, newdata=test2)

?predict

predictions <- predict(modelFit, newdata=testing)
predictions

```

Random forest

```{r}

modFit <- train(classe~., data=train2, method="rf", prox=T)
modFit

getTree(modFit$finalModel, k=2)

pred <- predict(modFit, test)
test$predRight <- pred==test$Species
table(pred, test$Species)



classeP <- classCenter(train2[,c(3,4)], train2$classe, modFit$finalModel$prox)
classeP <- as.data.frame(classeP)
classeP$classe <- rownames(classeP)
classeP

p <- qplot(Petal.Width, Petal.Length, col=Species, data=train)

p + geom_point(aes(x=Petal.Width, y=Petal.Length, col=Species), size=5, shape=4, data=irisP)


pred <- predict(modFit, test)
test$predRight <- pred==test$Species
table(pred, test$Species)